%%
%% This is acknowledgments.tex.
%% A content TeX-file of the Ph.D. thesis of Tom M. Ragonneau.
%% It follows the requirements of The Hong Kong Polytechnic University.
%%
%% It may be distributed and/or modified under the conditions of the
%% LaTeX Project Public License, either version 1.3 of this license or
%% (at your option) any later version. The latest version of this
%% license is in
%%     http://www.latex-project.org/lppl.txt
%% and version 1.3 or later is part of all distributions of LaTeX
%% version 2005/12/01 or later.
%%
%% This work has the LPPL maintenance status 'maintained'.
%%
%% The Current Maintainer of this work is Tom M. Ragonneau.
%%
\chapter*{Acknowledgments}
\label{ch:ack}
\addcontentsline{toc}{chapter}{\nameref*{ch:ack}}

\begin{itemize}
    \item The Research Grants Council of Hong Kong, under the aegis of the University Grants Committee of Hong Kong, for providing financial support under the Hong Kong Ph.D. Fellowship Scheme.
    \item The Hong Kong Polytechnic University for supporting the research study.
    \item The Department of Applied Mathematics of The Hong Kong Polytechnic University for providing the necessary equipments and research facilities.
    \item Dr.\ Zaikun Zhang (張在坤博士) and Prof.\ Xiaojun Chen (陳小君教授) from the Department of Applied Mathematics of The Hong Kong Polytechnic University for supervising the research study.
    \item Prof.\ Serge Gratton from the Department of Computational and Applied Mathematics of Toulouse INP/ENSEEIHT for sharing beneficial commentaries on the research project.
    \item Prof.\ Ya-xiang Yuan (袁亞湘教授) for his everlasting encouragement and support.
    \item My parents and my sister for their wholesome counsel, wise guidance, and sympathetic ears during the research study.
    \item My friends for being always by my side, providing invigorating discussions along with resourceful distractions.
\end{itemize}
